\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses

### 2015-04-20 doc/sw-system.tex
Working on Anu's comments, up to Section 3 #112

\title[\SW]
{\SW: Crowd-sourcing the Discovery of Gravitational Lenses}

@@

We describe \SW, a novel gravitaional lens discovery service that yields samples
of high purity and completeness through crowd-sourced visual inspection.
Carefully produced colour composite images are displayed to volunteers via a
classification interface which records their estimates of the positions of
candidate lensed features. Simulated lenses, and expert-classified images which lack lenses,
are inserted into the image stream at random intervals; this training set is
used to give the volunteers feedback on their performance, and to estimate a
dynamically-updated probability for any given image to contain a lens. Low
probability systems are retired from the site periodically, concentrating the
sample towards a set of candidates. Having divided the 171 square degree
Canada-France-Hawaii Telescope Legacy Survey (\cfhtls)
imaging survey into some 430,000 overlapping 80 by 80 arcsecond tiles and
displayed them on the site, we were joined by around 37,000 volunteers who
contributed 10 million image classifications over the course of 8 months. The
sample was reduced to 3300 \StageOne candidates; these were then refined to
yield a sample of 90 \StageTwo detections, a sample we expect to be over 90\%
complete and 20\% pure at a threshold of 95\% classification probability. We
comment on the scalability of the \SW system to the wide field survey era.

@@

For the more interesting subjects, the CI offers two features that  enable
further investigation of the subjects. The first is the ``Quick Dashboard''
(QD) a more advanced image viewer. This allows the viewer to compare three
different contrast and colour balance settings, to help bring out subtle
features, and to pan and zoom in on interesting regions of the image to assess
small features. Markers can be placed in the Quick Dashboard just the same as
in the main CI image viewer. The second is a link to that subject's page in
the project discussion forum, \texttt{http://talk.spacewarps.org} (known as
``Talk''). Here, volunteers can discuss the features they have seen either
before they submit their classification, or after, if they ``favourite'' the
subject. There is no ``back'' button: each volunteer may only classify a given
subject once. However, the presence of an option to see what others think
about any given subject before submitting your own classification means that
the classifications may not be strictly independent; the advantage of this
system is that volunteers can learn from others what constitutes a good lens
candidate. This is not, however, the primary educational resource; we describe
the explicit training that we provide for the volunteers in the next section.
The FITS images of the subject can be further explored with Zooniverse
``Dashboard Tools,'' using a more powerful image viewer that enables dynamic
variation of the colour balance and contrast (stretch), and for any given view
to be shared via unique URL back to Talk. The FITS image viewing in both the
Quick Dashboard and the Dashboard Tools are enabled by the javascript library
\fitsjs \citep{fitsjs}.

@@

When a volunteer begins classifying, they are initially shown training images at
a frequency of two in five; these subjects are drawn at random from a pool
containing equal numbers of sims and duds, such that that volunteer never sees a given image more than once. As the number of classifications made by a
volunteer increases, this frequency is decreased, to
$2/(5\times2^{(\textrm{int}(N_c/20)+1)/2})$ ($\approx 0.3$ for the second 20
subjects, $0.2$ for the third 20 subjects, and so on).

@@

\subsubsection{Spotter's Guide}

The instant feedback provides real-time educational responses to the
volunteers as they start classifying; as well as this dynamic system, \SW
provides a static reference work for volunteers to consult when in doubt about
how to perform the task. This ``Spotter's Guide'' is a set of webpages showing
example lenses, both real and simulated, and also some common false positives,
drawn from the pool of survey images. The non-lenses were identified by three
of us (AV, AM and PM) while inspecting a small set of survey images in order
to define the ``dud'' training images. For easy reference, the lenses are
divided by type (for example, ``lensed galaxies,'' ``lensed quasars'' and
``cluster lenses''), as are the false positives (for example, ``Rings and
Spirals,'' ``Mergers,'' ``Artifacts'' and so on). The example images are
accompanied by explanatory text. The Spotter's Guide is reached via a button
on the left hand side, or the hyperlinked thumbnail images of the ``Quick
Reference'' provided on the right hand side, of the classification interface.

Most of the text of the Spotter's Guide focuses on what lenses do or don't do;
the website ``Science page'' contains a very brief introduction to how
gravitational lenses work, which is fleshed out a little on the ``FAQ'' page.
This also contains answers to frequently asked questions about the interface
and the task set.

@@

The \StageOne training set was chosen to be quite clear cut, in order to err on
the side of inclusivity, and so ensure high completeness. For the training duds
we selected several hundred images at random, and three of us (AM, with AV and
PM) inspected them and discarded anything that could be considered a lens
candidate: this meant that objects that looks similar to lenses, such as galaxy
mergers, tidal tails and spiral arms, pairs of blue stars and so on, were
specifically excluded from the training set, and so we expect some of those
types of object to appear in the \StageOne candidate list. The training sims for
\StageOne were also selected to be relatively straightforward to spot, as we sought a balance between sensitivity to faint or difficult lenses, and the need to educate the volunteers in what gravitational lenses look like. We return to this balance below.

@@

The design of the \StageOne classification task, and its training set, should
lead to sample of lens candidates that has high completeness but may have low
purity. To refine this sample to higher purity, we need to reject more
non-lenses, which means providing the volunteers with a more realistic and
challenging training set as they re-classify it. The more demanding \StageTwo
training set was generated as follows. The \StageTwo duds were selected from the
image pool by at least one of three of us  (AV, AM and PM) as being a potential
false positive, while the \StageTwo sims were chosen to be a subset of \StageOne
sims, none of which were deemed ``obvious'' by the same expert classifiers.
Figures~\ref{fig:training-gallery:sims} and~\ref{fig:training-gallery:duds}
show some example images from the resulting \cfhtls \StageTwo training set.

@@

We also attempted to encourage discernment by changing the look and feel of
the app, slowing down the arrival of new images, and switching the background
colour to bright orange to make it clear that a different task was being set.
The frequency with which training images were shown was fixed at 1 in 3.
Finally, the Spotter's Guide was upgraded to include more examples
of various possible false positives, divided into sub-classes.

@@

Each classification made is logged in a database, storing subject IDs,
(anonymous) volunteer IDs, a timestamp and the classification results (the image
pixel-coordinate positions of every marker placed).  The ``category'' of subject --
whether it is a ``training'' subject (a simulated lens or a known non-lens) or a
``test'' subject (an unseen image drawn from the survey) -- is also recorded.
For training subjects, we also store the ``kind'' of the subject as
a lens (``sim''), or a non-lens (``dud''), and also the ``flavour'' of lens
object if one is present in the image (``lensed galaxy'', ``lensed quasar'' or
``cluster lens'').  This classification is used to provide instant feedback, but
is also the basic measurement used in a probabilistic classification of every
subject based on all image views to date.

>>>>

\title[\SW]
{\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses}

@@

We describe \SW, a novel gravitaional lens discovery service that yields samples
of high purity and completeness through crowd-sourced visual inspection.
Carefully produced colour composite images are displayed to volunteers via a
classification interface which records their estimates of the positions of
candidate lensed features. Simulated lenses, and expert-classified non-lenses,
are inserted into the image stream at random intervals; this training set is
used to give the volunteers feedback on their performance, and to estimate a
dynamically-updated probability for any given image to contain a lens. Low
probability systems are retired from the site periodically, concentrating the
sample towards a set of candidates. Having divided the 171 square degree
Canada-France-Hawaii Telescope Legacy Survey (\cfhtls)
imaging survey into some 430,000 overlapping 80 by 80 arcsecond tiles and
displayed them on the site, we were joined by around 37,000 volunteers who
contributed 10 million image classifications over the course of 8 months. The
sample was reduced to 3300 \StageOne candidates; these were then refined to
yield a sample of 90 \StageTwo detections, a sample we expect to be over 90\%
complete and 20\% pure at a threshold of 95\% classification probability. We
comment on the scalability of the \SW system to the wide field survey era.

@@

For the more interesting subjects, the CI offers two features that  enable
further investigation of the subjects. The first is the ``Quick Dashboard''
(QD) a more advanced image viewer. This allows the viewer to compare three
different contrast and colour balance settings, to help bring out subtle
features, and to pan and zoom in on interesting regions of the image to assess
small features. Markers can be placed in the Quick Dashboard just the same as
in the main CI image viewer. The second is a link to that subject's page in
the project discussion forum, \texttt{http://talk.spacewarps.org} (known as
``Talk''). Here, volunteers can discuss the features they have seen either
before they submit their classification, or after, if they ``favorite'' the
subject. There is no ``back'' button: each volunteer may only classify a given
subject once. However, the presence of an option to see what others think
about any given subject before submitting your own classification means that
the classifications may not be strictly independent; the advantage of this
system is that volunteers can learn from others what constitutes a good lens
candidate. This is not, however, the primary educational resource; we describe
the explicit training that we provide for the volunteers in the next section.
The FITS images of the subject can be further explored with Zooniverse
``Dashboard Tools,'' using a more powerful image viewer that enables dynamic
variation of the colour balance and contrast (stretch), and for any given view
to be shared via unique URL back to Talk. The FITS image viewing in both the
Quick Dashboard and the Dashboard Tools are enabled by the javascript library
\fitsjs \citep{fitsjs}.

@@

When a volunteer begins classifying, they are initially shown training images at
a mean frequency of two in five. Subjects are drawn at random from a pool
consisting of (at first) 20\% sims, 20\% duds, and 60\% test images,
and such that no volunteer ever
sees a given image more than once. As the number of classifications made by the
volunteer increases, the training frequency is decreased from 40\% to
$2/(5\times2^{(\textrm{int}(N_c/20)+1)/2})$ ($\approx 30\%$ for the second 20
subjects, $20\%$ for the third 20 subjects, and so on).

@@

\subsubsection{Spotter's Guide}
\label{sec:design:training:guide}

The instant feedback provides real-time educational responses to the
volunteers as they start classifying; as well as this dynamic system, \SW
provides a static reference work for volunteers to consult when in doubt about
how to perform the task. This ``Spotter's Guide'' is a set of webpages showing
example lenses, both real and simulated, and also some common false positives,
drawn from the pool of survey images. The non-lenses were identified by three
of us (AV, AM and PM) while inspecting a small set of survey images in order
to define the ``dud'' training images (\Sref{sec:design:stages}). For easy reference, the lenses are
divided by type (for example, ``lensed galaxies,'' ``lensed quasars'' and
``cluster lenses''), as are the false positives (for example, ``Rings and
Spirals,'' ``Mergers,'' ``Artifacts'' and so on). The example images are
accompanied by explanatory text. The Spotter's Guide is reached via a button
on the left hand side, or the hyperlinked thumbnail images of the ``Quick
Reference'' provided on the right hand side, of the classification interface.

Most of the text of the Spotter's Guide focuses on what kinds of image features
gravitational lenses do or don't produce;
the website ``Science page'' contains a very brief introduction to how
gravitational lensing works, which is fleshed out a little on the ``FAQ'' page.
The FAQ also contains answers to frequently asked questions about the interface
and the task set.

@@

The \StageOne training set was chosen to be quite clear cut, in order to err on
the side of inclusivity, and so ensure high completeness. We selected the
training duds by visual inspection, discarding anything that could be considered
a lens candidate (see \Sref{sec:design:training:guide}): this meant that objects
that look similar to lenses, such as galaxy mergers, tidal tails and spiral
arms, pairs of blue stars and so on, were specifically excluded from the
training set, and so we expect some of those types of object to appear in the
\StageOne candidate list. The training sims for \StageOne were also selected to
be relatively straightforward to spot, as we sought a balance between
sensitivity to faint or difficult lenses, and the need to educate the volunteers
in what gravitational lenses look like. We return to this balance below.

@@

The design of the \StageOne classification task, and its training set, should
lead to sample of lens candidates that has high completeness but may have low
purity. To refine this sample to higher purity, we need to reject more
non-lenses, which means providing the volunteers with a more realistic and
challenging training set as they re-classify it. The more demanding \StageTwo
training set was generated as follows. The \StageTwo duds were again selected
from the image pool by visual inspection as being a potential false positive,
while the \StageTwo sims were chosen to be a subset of the \StageOne sims, none
of which were deemed ``obvious'' by the same expert classifiers. This meant that
the \StageTwo sims were had fainter and less well-separated image features than
in the \StageOne training set.  Figures~\ref{fig:training-gallery:sims}
and~\ref{fig:training-gallery:duds} show some example images from the resulting
\cfhtls \StageTwo training set.

@@

We also attempted to encourage discernment by changing the look and feel of the
app, slowing down the arrival of new images, and switching the background colour
to bright orange to make it clear that a different task was being set. The
frequency with which training images were shown was fixed at 1 in 3. Finally,
the Spotter's Guide was upgraded to include more examples of various possible
false positives, divided into sub-classes. We did not retire any subjects during
\StageTwo classification, instead continuing to accumulate classifications
through till the end of a fixed 4-week time period.

@@

Each classification made is logged in a database, storing subject IDs,
(anonymous) volunteer IDs, a timestamp and the classification results (the image
pixel-coordinate positions of every marker placed).  The ``category'' of subject --
whether it is a ``training'' subject (a simulated lens or a known non-lens) or a
``test'' subject (an unseen image drawn from the survey) -- is also recorded.
For training subjects, we also store the ``kind'' of the subject as
a lens (``sim''), or a non-lens (``dud''), and also the ``flavor'' of lens
object if one is present in the image (``lensed galaxy'', ``lensed quasar'' or
``cluster lens'').  This classification is used to provide instant feedback, but
is also the basic measurement used in a probabilistic classification of every
subject based on all image views to date.
